Section_8_Team1_Lab2_Report

### 2021-08-04 reports/Draft_Report.Rmd
updated EDA visuals

```{r setup, message=FALSE}
install.packages("table1")

#we may not need all of these
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(dplyr)
library(table1)
library(ggally)
theme_set(theme_minimal())
knitr::opts_chunk$set(echo = TRUE)

@@

```{r dataframe, echo=FALSE}
df <- read.csv('/home/jovyan/r_bridge/code/lab-2-lab2_section8_jal/data/interim/covid_trips_interim_model3.csv')
head(df)

@@

To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas. For the analysis, ### responses were analyzed as shown in Table 1. Table 2 depicts the county type breakdown for rural or urban classification as well.

@@

step_num = c(1,2,3)

number_of_samples = c(nrow(df),
                      nrow(#newdf1),
                      nrow(#newdf2))                      
samples_removed = c(0,
                    nrow(df)-nrow(#newdf1),
                    nrow(#newdf1)-nrow(#newdf2))

reasoning = c("Original Dataset",
              "Removing XYZ",
              "Removing 123")

@@

```{r EDA}
#Example Table
table1::label(df$total_pop) <- "TX County Population"
table1::label(df$sum_cases) <- "TX County Covid-19 Cases"

table1::table1(~total_pop + sum_cases | state, data = df)

#Short vs long trip: rate of short trips is higher than rate of long trips in relation to # of covid cases
ggplot(data = df) + aes(x = log(sum_cases), y = log(long_trip)) + geom_point()

ggplot(data = df) + aes(x = log(sum_cases), y = log(short_trip)) + geom_point()

df %>% 
  mutate(
    total_pop = log(total_pop),
    cases = log(cases+1),
    deaths = log(deaths+1),
    long_trip = log(long_trip),
    density = log(density)
  ) %>%
  select(long_trip, cases, deaths, median_income, unempoy_rate, density, pct_lt17) %>% 
  GGally::ggpairs(.,
               legend = 1,
               mapping = ggplot2::aes(colour=as.factor(df$is_rural)))

>>>>

```{r setup, message=FALSE}
install.packages("table1")
install.packages("GGally")

#we may not need all of these
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(dplyr)
library(table1)
library(GGally)
theme_set(theme_minimal())
knitr::opts_chunk$set(echo = TRUE)

@@

```{r dataframe, echo=FALSE}
df <- read.csv('~/lab-2-lab2_section8_jal/data/interim/covid_trips_interim_model3.csv')
#head(df)
```

Figure 1 below depicts the distribution and correlation of each variable in our dataset by county type (where '0' indicates an urban county). From Figure 1, it can be observed that some of the variables are not distributed normally, are skewed,  and thus require transformations. More specifically, these variables are: Total Population, Total Cases, Total Deaths, Population Density. The outcome variable, Long Distance, will also require transformation. Upon further analysis, logarithmic transformations were chosen to normalize the distribution and minimize skewness.

```{r EDA}
#Example Table
df$rural_cat <- ifelse(df$is_rural == 1, "Rural Counties", "Urban Counties")
table1::label(df$total_pop) <- "County Population"
table1::label(df$density) <- "County Population Density"
table1::label(df$cases) <- "County Covid-19 Cases"
table1::label(df$unempoy_rate) <- "County Unemployment Rate"
table1::label(df$median_income) <- "County Median Income ($)"
table1::label(df$pct_lt17) <- "County Age < 17 years (%)"
table1::label(df$pct_gt65) <- "County Age > 65 years (%)"
# table1::label(df$is_rural) <- "County Type"
table1::table1(~total_pop + density + cases + unempoy_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")

#Short vs long trip: rate of short trips is higher than rate of long trips in relation to # of covid cases
ggplot(data = df) + aes(x = log(cases), y = log(long_trip)) + geom_point()
#ggplot(data = df) + aes(x = log(cases), y = log(short_trip)) + geom_point()

Rural_County=as.factor(df$is_rural)

df %>% 
  mutate(
    total_pop = log(total_pop),
    cases = log(cases+1),
    deaths = log(deaths+1),
    long_trip = log(long_trip),
    density = log(density)
  ) %>%
  select(long_trip, cases, deaths, median_income, unempoy_rate, density, pct_lt17) %>% 
  GGally::ggpairs(.,
               legend = 1, title = "Figure 1: Variable Statistics",
               mapping = ggplot2::aes(colour=Rural_County))

@@

To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas. For the analysis, 230 responses were analyzed as shown in Table 1. Table 2 depicts the county type breakdown for rural or urban classification as well.

@@

df2 <- df

df2$log_pop <- log10(df2$total_pop)
df2$log_cases <-log10(df2$cases)
df2$log_deaths <-log10(df2$deaths)
df2$log_long_trip <-log10(df2$long_trip)

step_num = c(1,2)

number_of_samples = c(nrow(df),
                      nrow(df2))
samples_removed = c(0,
                    nrow(df)-nrow(df2))

reasoning = c("Original Dataset",
              "Logarithmic Transformations")

@@
